WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-11-19 ms.tex
Remove Larry's Caltech e-mail address
Increment DCC number
Revert "Remove Larry's Caltech e-mail address"
This reverts commit c33eb20af5eeffb6ff33fac783c06f2d5c0570b6.
Remove some comments that have been addressed
Remove qualifications about phase convention
This is, in fact, the phase convention used by FINDCHIRP and LALSimulation.

% Establish and describe the first rapid and tractable solution of the GW position reconstruction problem.
% Use and abuse of Fisher matrix: use it as a clue.
% Generality: replace GW antenna pattern and GW polarization model with any other kind of receiver.
% Demonstration of a practical Bayesian analysis that does not use an MCMC method.

@@

Assuming a nonprecessing circular orbit, we can write the \ac{GW} signal received by any detector as a linear combination of two basis waveforms, $H_0$ and $H_{\pi/2}$ \citep{PhysRevD.83.084002}. $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the sine and cosine functions, being nearly orthogonal and out of phase by ${\pi/2}$ at all frequencies. If $H_0$ and $H_{\pi/2}$ are Fourier transforms of real functions, then $H_0(\omega) = H_0^*(-\omega)$ and $H_{\pi/2}(\omega) = H_{\pi/2}^*(-\omega)$, and we can write (assuming an arbitrary phase convention)
%
\begin{equation}
    H_{\pi/2}(\omega) = H_0(\omega) \cdot
    \begin{cases}
        -i & \text{if } \omega \geq 0 \\
        i & \text{if } \omega < 0
    \end{cases}.
\end{equation}
%
For brevity, we define $H \equiv H_{0}$ and write all subsequent equations in terms of the $H$ basis vector alone. Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters, $\bm\theta_\mathrm{ex}$, into a few coefficients, and all dependence on the intrinsic parameters, $\bm\theta_\mathrm{in}$, into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
%
\begin{multline}\label{eq:full-signal-model}
    X_i(\omega; \bm\theta) = e^{-i \omega (t_\oplus - \mathbf{d}_i \cdot \mathbf{n})}
    \frac{r_{1,i}}{r}
    e^{2 i \phi_c}
    \\
    \left[
    \frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{\zeta\right\} - i
    \left(\cos\iota\right) \Im \left\{\zeta\right\}
    \right]
    H(\omega; \bm\theta_\mathrm{in})
\end{multline}
%
for $\omega \geq 0$, where
%
\begin{equation}
    \zeta = e^{-2 i \psi} \left(
    F_{+,i}(\alpha, \delta, t_\oplus) +
    i F_{\times,i}(\alpha, \delta, t_\oplus)
    \right).
\end{equation}
%
The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}}^2 \leq 1$. They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event and are presented in~\cite{ExcessPower}. In a coordinate system with the $x$ and $y$ axes aligned with the arms of a detector, the antenna pattern is given in spherical polar coordinates as
%
\begin{align}
    F_+ &= -\frac{1}{2}(1 + \cos^2 \theta) \cos{2\phi}, \\
    F_\times &= -\cos\theta \sin{2\phi}.
\end{align}
%
The unit vector $\mathbf{d}_i$ represents the position of detector $i$ in units of light travel time.\footnote{When considering transient \ac{GW} sources such as those that we are concerned with in this thesis, the origin of the coordinate system is usually taken to be the geocenter. For long\nobreakdashes-duration signals such as those from statically deformed neutron stars, the solar system barycenter is a more natural choice.} The vector $\mathbf{n}$ is the direction of the source. The negative sign in the dot product $-\mathbf{d}_i \cdot \mathbf{n}$ is present because the direction of travel of the \ac{GW} signal is opposite to that of its sky location. The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms):
%
\begin{equation}\label{eq:horizon}
r_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H(\omega; \boldsymbol \theta_\mathrm{in})\right|^2}{S_i(\omega)} \,d\omega.
\end{equation}

@@

\begin{acknowledgements}
This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v3.

>>>>

@@

Assuming a nonprecessing circular orbit, we can write the \ac{GW} signal received by any detector as a linear combination of two basis waveforms, $H_0$ and $H_{\pi/2}$ \citep{PhysRevD.83.084002}. $H_0$ and $H_{\pi/2}$ are approximately ``in quadrature'' in the same sense as the cosine and sine functions, being orthogonal and out of phase by ${\pi/2}$ at all frequencies. If $H_0$ and $H_{\pi/2}$ are Fourier transforms of real functions, then $H_0(\omega) = H_0^*(-\omega)$ and $H_{\pi/2}(\omega) = H_{\pi/2}^*(-\omega)$, and we can write
%
\begin{equation}
    H_{\pi/2}(\omega) = H_0(\omega) \cdot
    \begin{cases}
        -i & \text{if } \omega \geq 0 \\
        i & \text{if } \omega < 0
    \end{cases}.
\end{equation}
%
For brevity, we define $H \equiv H_{0}$ and write all subsequent equations in terms of the $H$ basis vector alone. Then, we can write the signal model in a way that isolates all dependence on the extrinsic parameters, $\bm\theta_\mathrm{ex}$, into a few coefficients, and all dependence on the intrinsic parameters, $\bm\theta_\mathrm{in}$, into the basis waveform, by taking the Fourier transform of Equation~(2.8) of \cite{PhysRevD.83.084002}:
%
\begin{multline}\label{eq:full-signal-model}
    X_i(\omega; \bm\theta) = e^{-i \omega (t_\oplus - \mathbf{d}_i \cdot \mathbf{n})}
    \frac{r_{1,i}}{r}
    e^{2 i \phi_c}
    \\
    \left[
    \frac{1}{2} \left(1 + \cos^2 \iota\right) \Re \left\{\zeta\right\} - i
    \left(\cos\iota\right) \Im \left\{\zeta\right\}
    \right]
    H(\omega; \bm\theta_\mathrm{in})
\end{multline}
%
for $\omega \geq 0$, where
%
\begin{equation}
    \zeta = e^{-2 i \psi} \left(
    F_{+,i}(\alpha, \delta, t_\oplus) +
    i F_{\times,i}(\alpha, \delta, t_\oplus)
    \right).
\end{equation}
%
The quantities $F_{+,i}$ and $F_{\times,i}$ are the dimensionless detector antenna factors, defined such that $0 \leq {F_{+,i}}^2 + {F_{\times,i}}^2 \leq 1$. They depend on the orientation of detector $i$ as well as the sky location and sidereal time of the event and are presented in~\cite{ExcessPower}. In a coordinate system with the $x$ and $y$ axes aligned with the arms of a detector, the antenna pattern is given in spherical polar coordinates as
%
\begin{align}
    F_+ &= -\frac{1}{2}(1 + \cos^2 \theta) \cos{2\phi}, \\
    F_\times &= -\cos\theta \sin{2\phi}.
\end{align}
%
The unit vector $\mathbf{d}_i$ represents the position of detector $i$ in units of light travel time.\footnote{When considering transient \ac{GW} sources such as those that we are concerned with in this thesis, the origin of the coordinate system is usually taken to be the geocenter. For long\nobreakdashes-duration signals such as those from statically deformed neutron stars, the solar system barycenter is a more natural choice.} The vector $\mathbf{n}$ is the direction of the source. The negative sign in the dot product $-\mathbf{d}_i \cdot \mathbf{n}$ is present because the direction of travel of the \ac{GW} signal is opposite to that of its sky location. The quantity $r_{1,i}$ is a fiducial distance at which detector $i$ would register SNR=1 for an optimally oriented binary (face\nobreakdashes-on, and in a direction perpendicular to the interferometer's arms):
%
\begin{equation}\label{eq:horizon}
r_{1,i} = 1 / \sigma_i, \qquad {\sigma_i}^2 = \int_0^\infty \frac{\left|H(\omega; \boldsymbol \theta_\mathrm{in})\right|^2}{S_i(\omega)} \,d\omega.
\end{equation}

@@

\begin{acknowledgements}
This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v4.
